---
title: "base plot"
author: Raphael Rehms
format: 
  revealjs:
    scrollable: true
    theme: solarized # moon
    transition: slide
    footer: "Control flows and programming"
    slide-number: c
    include-after-body: clean_title_page.html
    fontsize: 25pt
    highlight-style: github
execute: 
  echo: true
  freeze: auto
editor: 
  markdown: 
    wrap: 72
---

# Visualization of Data

## Graphics general
- There are several kind of graphics available in R. E.g.

    - ```plot()```;  ```hist()```;  ```boxplot()```;

- The title of a pplot is set with the additional argument
    - ```main = 'Main title'```

- The axes names by the additional arguments:

    - ```xlab = 'x'``` for the x-axis; ```ylab = 'y'``` for the y-axis

    - Color can be set by argument col (e.g., ```col='red'```, ```col='grey'```, . . . )




## Plot
- Standard scatter or line plot

- Needs the value for x-axis and the y-axis

- different types can be chosen (via argument ```type```):

    - 'p' for points (default); 'l' for lines

    - A lot of styling options

    - See help for more details



## Scatter plot
A basic scatter plot is created with the basic plot function
```{r}
#| fig-align: "center"

x <- -5:5
y <- (-5:5)^2
plot(x, y, main="Graphic 1")
```


## Line plot
- A basic line plot is also created with the basic plot function (with additional type)

```{r}
#| fig-align: "center"

plot(y ~ x, main="Graphic 2", type="l")
```




## Histogram
- Command ```hist()```

- Use the argument breaks to have more or less bars;

- Set option freq = TRUE for frequencies, freq = FALSE for densities;



## Histogram set color
- Command ```hist()```

```{r}
#| fig-align: "center"

attach(sleep)
hist(extra, main="Graphic 3", col="grey")
```



## Histogram set breaks 
- A basic histogram with a custom number of breaks

```{r}
#| fig-align: "center"

attach(sleep)
hist(extra, main="Graphic 4", breaks=20, col=2)
```

## Boxplot

- Command ```boxplot()```;

- Easy to plot with different groups;

- Parts of boxplot:
    - median, 
    - 1^st^ and 3^rd^ quartile
    - outliers



## Single boxplot
- A basic boxplot 

```{r}
#| fig-align: "center"

attach(sleep)
boxplot(extra, main="Graphic title")
```


## Multiple boxplots 

A boxplot with formula operator to make a boxplot for each group

```{r}
#| fig-align: "center"

boxplot(extra ~ group, main="Boxplot for each group", ylab="extra", xlab="group")
```


# Exercise

