---
title: "Models in R"
author: Raphael Rehms und Jonathan Christ
format: 
  revealjs:
    scrollable: true
    theme: solarized # moon
    transition: slide
    footer: "Models in R"
    slide-number: c
    include-after-body: clean_title_page.html
    fontsize: 25pt
    highlight-style: github
execute: 
  echo: true
  freeze: auto
editor: 
  markdown: 
    wrap: 72
---

# Models in R

## Linear model

$Y =\beta_0+\beta_1 * X_1+ \beta_2 * X_2+...+\epsilon$ ,with
$\epsilon \sim N(0,\sigma^2)$

R command: `lm(formula, data, ...)`

examples of `formula`:

-   `y~age+sex`

-   `y~I(log(age)) + as.factor(sex)`

-   see `?formula` for more details

Note: `data` must be a data.frame.

Example:

```{r}
library(MASS)
data(cats)
with(cats, plot(Bwt, Hwt, xlab="Body Weight (kg)",
        ylab="Heart Weight (g)",
         main="Heart Weight vs. Body Weight of Cats"))

with(cats, cor.test(Bwt, Hwt))
lm(Hwt ~ Bwt,data=cats)
```

## Function summary()

-   The output of `summary()` provides an overview on the model:

```{r}
model <- lm(Hwt ~ Bwt,data=cats)
summary(model)
```

## Graphical diagnosis

-   Plot the regression line

```{r}
plot(cats$Bwt, cats$Hwt, xlab="Body Weight (kg)",
        ylab="Heart Weight (g)",
         main="Heart Weight vs. Body Weight of Cats")
yhat <- coef(model)[1] + coef(model)[2]*cats$Bwt

lines(x=cats$Bwt,y=yhat,col=2)

```

## Graphical diagnosis

Use the function `plot()` for diagnosis graphics:

```{r}
plot(yhat,residuals(model))

qqnorm(residuals(model))
qqline(residuals(model))

plot(model)
```

## Model selection

-   In R there are pre-built functions which perform automatic model
    selection, for example, `stepAIC()` in the package MASS

```{r,eval=FALSE}
stepAIC(object, scope, scale = 0, direction = c("both", "backward", "forward"), trace = 1, keep = NULL, steps = 1000, k = 2, ...);
```

-   By default, it uses AIC as stopping criterion (argument `k=2`)

-   setting `k=log(n)` we can use BIC. Example:

```{r}
library(MASS)
model <- lm(Hwt ~ Bwt+Sex,data=cats)
model.new<-stepAIC(model)
```

## Prediction

-   The function `predict()` computes the predicted value of the
    response for a new observation.

```{r, eval=FALSE}
predict(object, newdata, se.fit = FALSE, interval = c("none", "confidence", "prediction"), level = 0.95, type = c("response", "terms"), ...)
```

in particular:

`object` must be replaced by the model fit on the data

`newdata` must be a data.frame with the new observation(s)

`se.fit` allows the computation of the standard error

setting `interval="prediction"`, we can compute the prediction interval,
with `level` (default `level = 0.95`);

- Example

```{r, eval=FALSE}
model <- lm(Hwt ~ Bwt,data=cats)
y.hat <- predict(model, newdata = cats, interval="prediction")
predict(model) == coef(model)[1] + coef(model)[2]*cats$Bwt
```

## Logistic model

$logit(\pi)=\beta_0+\beta_1 * X_1+ \beta_2 * X_2+...$ ,with
$\pi=Pr(Y=1|X)$

-   is a special case of generalized linear model

```{r,eval=FALSE}
glm(formula, family = gaussian, data, ...)
```

-   to fit a logistic model, the argument `family` must be set equal to
    binomial.

Example:

```{r}
load(file="myNhanes.RData")
glm(cancer_ever~workpollut,family=binomial,data=mydata)
```

## Like for the linear model...

-   we can have an overview on the model through the function
    `summary()`:

```{r}
summary(glm(cancer_ever~workpollut,family=binomial, data=mydata))
```

## Model Selection 

- we can perform model selection (e.g., based on AIC) through the
function `stepAIC()`

- we can use the function predict() to predict


```{r, eval=FALSE}
stepAIC(glm(mydata$cancer_ever~mydata$workpollut,family=binomial))

predict(glm(cancer_ever~workpollut,family=binomial, data=mydata),newdata=mydata)
```
