---
title: "Normal Distribution"
format: 
  revealjs:
    scrollable: true
    theme: solarized # moon
    transition: slide
    footer: "dplyr and data.table"
    slide-number: c
    include-after-body: clean_title_page.html
    fontsize: 25pt
    highlight-style: github
execute:
  echo: true
  warning: false
  message: false
  freeze: auto
  fig-align: "center"
  fig-width: 6
  fig-height: 4
editor: 
  markdown: 
    wrap: 72
---

# Normal Distribution


## Mathematical notation
$$
f(x; \mu, \sigma) = \frac{1}{\sigma \sqrt{2\pi}} e^{-\frac{(x - \mu)^2}{2\sigma^2}}
$$

Where:

- $x$ is the random variable.

- $\mu$ is the mean of the distribution.

- $\sigma$ is the standard deviation of the distribution.




## Sample from Normal Distribution

- generate a sample from a Gaussian distribution:

  ```rnorm(n, mean = 0, sd = 1)```

**Example:**
```{r}
#| layout: [[100]]
#| fig-align: "center"

y <- rnorm(n = 100000, mean = 5, sd = sqrt(2))
hist(y, freq = F, ylim = c(0, 0.3))
curve(dnorm(x, mean = 5, sd = sqrt(2)), col = 2, add = T)
```
<br>



## Density function (pdf) of Normal Distribution

Calculate pdf of Normal distribution:

-  ```dnorm(x, mean = 0, sd = 1)```

```{r}
#| layout: [[100]]
#| fig-align: "center"

y <- rnorm(n = 100000, mean = 0,sd = 1)
hist(y, freq=F, ylim = c(0, 0.4),breaks = 100)
dnorm(-1)
hist(y, freq = F, ylim = c(0, 0.4), breaks = 100)
abline(v = -1, lty = 2)
abline(h = dnorm(-1), lty=2)
```


## Cumullative Density function (cdf) of Normal Distribution


Calculate pdf of Normal distribution:

 - ```pnorm(q, mean = 0, sd = 1)```

```{r,warning=F,message=FALSE}
#| fig-align: "center"
#| layout: [[100]]

library(tigerstats)

pnorm(-1)
pnormGC(-1, region = "below", graph = T)
```


## Cumullative Density function (cdf) of Normal Distribution

```{r,warning=F,message=FALSE}
#| fig-align: "center"
#| layout: [[100]]

library(tigerstats)

plot(ecdf(y),main = "Empirical Cumulative Distribution Function")
abline(v= quantile(ecdf(y),0.158655254), lty = 2)
abline(h= pnorm(-1), lty = 2)
text(x = -0.15, y = 0.2,labels = "(-1,0.159)")
```





## Quantiles of Normal Distribution

 - ```qnorm(p, mean = 0, sd = 1)```

```{r}
#| fig-align: "center"
#| layout: [[100]]

qnorm(0.158655254)
quantile(ecdf(y),0.158655254)
plot(ecdf(y), main = "Empirical Cumulative Distribution Function")
abline(v = quantile(ecdf(y), 0.158655254), lty = 2)
abline(h = pnorm(-1), lty = 2)
text(x = -0.15, y = 0.2,labels = "(-1,0.159)")
```
<br>


# EXAMPLES

## 
Consider $X \sim N(0,1)$. It is very easy to compute the following probabilities with R:

- $P(X \le 0.89)$
```{r}
#| fig-align: "center"
#| layout: [[100]]

pnorm(0.89)
```

```{r}
#| fig-align: "center"

pnormGC(0.89, region="below",graph=T)
```


## 
- $P(X \ge 1.21)$
```{r}

#| fig-align: "center"
#| layout: [[100]]

1-pnorm(1.21)
pnormGC(1.21, region="above",graph=T)
```


## 
- $P(-1.02 \le X \le 1.98)$
```{r}
#| fig-align: "center"

pnorm(1.98)-pnorm(-1.02)
pnormGC(c(-1.02,1.98), region="between",graph=T)
```


## 
- $P(\mid X \mid \le 0.92)$
```{r}
#| fig-align: "center"
#| layout: [[100]]

pnorm(0.92)-pnorm(-0.92)
pnormGC(c(-0.92,0.92), region="between",graph=T)
```


## 
- $P(\mid X \mid \ge 1.11)$
```{r}
#| fig-align: "center"

(1-pnorm(1.11))+pnorm(-1.11)
pnormGC(c(-1.11,1.11), region="outside",graph=T)
```

<br><br>

## OTHER DISTRIBUTIONS
We can use other distributions like the Normal distribution, for example - binomial: rbinom, dbinom, pbinom, qbinom

- uniform: runif, dunif, punif, qunif

- Student's t: rt, dt, pt, qt

- ...

