---
title: "Tests in R"
author: Raphael Rehms und Jonathan Christ
format: 
  revealjs:
    scrollable: true
    theme: solarized # moon
    transition: slide
    footer: "Tests in R"
    slide-number: c
    include-after-body: clean_title_page.html
    fontsize: 25pt
    highlight-style: github
execute: 
  echo: true
  freeze: auto
editor: 
  markdown: 
    wrap: 72
---

# Tests in R

# Prelude

## CONFIDENCE INTERVALS

-   it is an interval estimate;

-   we construct an interval following a procedure that, if applied in a
    large number of replications of the experiment, gives intervals
    which contain the true value $1 - \alpha$ of the time;

-   $1 - \alpha$ is called confidence level;

-   usual values are 0.90, 0.95 and 0.99.

## EXAMPLES of CONFIDENCE INTERVALS

-   true mean, gaussian distribution, known variance:

```{r}
x <- rnorm(100,mean=1,sd=2)
mean.x <- mean(x)
low <- mean.x - qnorm(0.975)*2/sqrt(100)
up <- mean.x + qnorm(0.975)*2/sqrt(100)
c(low,up)
```

-   true mean, gaussian distribution, unknown variance:

```{r}
sd.x <- sd(x)
low <- mean.x - qt(0.975, df=99)*sd.x/sqrt(100)
up <- mean.x + qt(0.975, df=99)*sd.x/sqrt(100)
c(low,up)
```

# Excercise 4 Task 3

## STUDENT'S ONE-SAMPLE TEST

-   $H_0 : \mu = \mu_0$

-   $H_1 :\mu \neq \mu_0$

`t.test(x, mu=`$\mu_0$`, alternative='two-sided')`

`t.test(x, mu=`$\mu_0$`, alternative='greater')`

`t.test(x, mu=`$\mu_0$`, alternative='less')`

-   default alternative is 'two-sided'

```{r}
x <- rnorm(100, mean=3, sd=1)
t.test(x, mu=1)
```

## Further examples

Further examples: greater, less

```{r}
t.test(x, mu=1, alternative="greater")
t.test(x, mu=1, alternative="less")
```

## TWO-SAMPLE T-TEST (UNEQUAL VARIANCE)

-   $Var[X] \neq Var[Y]$

-   $H_0: \mu_X = \mu_Y$

`t.test(x, y, alternative=c('two.sided', 'less','greater'))`

Example:

```{r}
x <- rnorm(100, mean=4, sd=5)
y <- rnorm(100, mean=2, sd=2)
t.test(x, y)
```

## TWO-SAMPLE T-TEST (EQUAL VARIANCE)

-   $Var[X] = Var[Y]$

-   $H_0: \mu_X = \mu_Y$

`t.test(x, y, var.equal=T,alternative=c('two.sided', 'less', 'greater'))`

Example:

```{r}
x <- rnorm(100, mean=4, sd=1)
y <- rnorm(100, mean=2, sd=1)
t.test(x, y, var.equal=T)
```

## F-TEST FOR COMPARING VARIANCES

-   $H_0: \sigma^2_X = \sigma^2_Y$

-   $H_1: \sigma^2_X \neq \sigma^2_Y$

`var.test(x, y, ratio = 1,alternative=c('two.sided', 'less', 'greater'),conf.level = 0.95)`

Example:

```{r}
x <- rnorm(80, mean=4, sd=5)
y <- rnorm(20, mean=2, sd=2)
var.test(x, y)
```

## PROPORTION TEST 1

-   $H_0: p = p_0$

-   $H_1: p \neq p_0$

`prop.test(x, n, p)`

-   x: number of successes

-   n: total number of trials

-   p: proportion to be tested.

-   Example:

```{r}
# Load preprocessed version of the data saved as .RData file
load(file="myNhanes.RData")
prop.test(sum(mydata$male), length(mydata$male), p=0.5)
```

## PROPORTION TEST 2

-   $H_0: p_1 = p_2$

-   $H_1: p_1 \neq p_2$

`prop.test(table(x, y))`

-   x: first (categorical) variable

-   y second (binary) variable

-   Example:

```{r}
# prop.test(table(educ, male))
tbl <- table(mydata$educ, mydata$male)
prop.test(tbl)
```

## Quick Summary

```{r,echo=F, message=F}
#setwd("/nfsmb/koll/le-lien/MSC_RcourseA_WS1617")
load(file="myNhanes.RData")
attach(mydata)
```

```{r,eval=F}
t.test(height[male==T],height[male==F])

var.test(height[male==T],height[male==F])
```

-   and a Z-test, testing the equality of proportions:

```{r,eval=F}
table(male,heartdis_ever,dnn=c("Male","Heart disease"))
prop.table(table(male,!heartdis_ever,dnn=c("Male","Heart disease")),1)
prop.test(table(male,!heartdis_ever))
```

## Chi-Square test

-   $H_0: \pi_1 = \pi_2 = \pi$

```{r,eval=F}
chisq.test(x, y = NULL, correct = TRUE, p =rep(1/length(x), length(x)), rescale.p = FALSE,
           simulate.p.value = FALSE, B = 2000)
```

-   you can pass to the function:
    -   a contingency table;
    -   directly two categorical vectors.

Examples:

```{r}
x <- c("M","F","M","F","M","M","F","F","M")
y <- c("Y","N","N","N","Y","Y","N","Y","N")
chisq.test(x,y)
```

or, with a contingency table,

```{r}
tab <- table(x,y)
chisq.test(tab)
```

## Fisher test

-   exact test

```{r,eval=F}
fisher.test(x, y = NULL, workspace = 200000,
    hybrid = FALSE, control = list(), or = 1,
    alternative = "two.sided", conf.int = TRUE,
    conf.level = 0.95, simulate.p.value = FALSE, B =
    2000)
```

-   you can pass to the function two factor vectors

Examples:

```{r}
fisher.test(x,y)
```

. . . or a contingency table

```{r}
tab <- table(x,y)
fisher.test(tab)
```

## McNemar test

-   symmetry of rows and columns;

```{r,eval=F}
mcnemar.test(x, y = NULL, correct = TRUE)
```

-   you can pass to the function two factor vectors

Examples:

```{r}
mcnemar.test(x,y)
```

. . . or a contingency table

```{r}
tab <- table(x,y)
mcnemar.test(tab)
```

# Excercise 4 Task 4

# Non-parametric Tests

## Sign test

-   $H_0 : median = md_0$

-   we need the package BSDA

```{r,eval=F}
SIGN.test(x, y = NULL, md = 0, alternative = "two.sided", conf.level = 0.95)
```

Example:

```{r, message=FALSE}
library(BSDA)
x <- rpois(100,5)
SIGN.test(x, md=5)
```

## Wilcoxon test

-   $H_0: median = md_0$

-   test based on ranks

```{r,eval=F}
wilcox.test(x, y = NULL, alternative = c("two.sided", "less", "greater"), mu = 0,
            paired = FALSE, exact = NULL, correct = TRUE, conf.int = FALSE,
            conf.level = 0.95, ...)
```

Example:

```{r}
x <- rpois(100,5)
wilcox.test(x,mu=5)
```

## Recall: Sign test vs. Wilcoxon test vs. t-test part 1

Sign test:

-   Applicable for ordinal or higher scale level

-   No distributional assumptions

-   Low power

Wilcoxon test:

-   Applicable for discrete or continuous quantitative data

-   Symmetrical distribution required

-   Intermediate power

T-test:

-   Normal distribution required

-   Highest power

## Compare all tests on the same data (we assume Normal distributed data!)

```{r}
set.seed(123)
x <- rnorm(20,0.5)
c(
  signTest = BSDA::SIGN.test(x, md = 0)$p.value,
  wilcoxonTest =wilcox.test(x,mu=0)$p.value,
  tTest = t.test(x, mu = 0)$p.value
)


```

## Mann-Whitney U test

-   $H_0: median_1 = median_2$

-   non-parametric test for independent samples

-   package coin (Why extra package and not using `wilcox.test`? -\>
    <https://stats.stackexchange.com/questions/31417/what-is-the-difference-between-wilcox-test-and-coinwilcox-test-in-r>)

```{r,eval=F}
wilcox_test(formula, data, subset = NULL, weights= NULL, ...)
```

-   needs a formula, e.g. y\~x.

Example hdl in NHANES data (not normal):

```{r}
hist(mydata$hdl[mydata$male == T], breaks = 40, col = "steelblue", freq = F, ylim=c(0,1.5))
hist(mydata$hdl[mydata$male == F], breaks = 40, col = "coral", add = T, freq = F)

coin::wilcox_test(hdl ~ as.factor(male), data = mydata)
wilcox.test(hdl ~ as.factor(male), data = mydata)

```

## Correlation

```{r,eval=F}
cor(X,Y)

cor(x, y = NULL, use = "everything", method =c("pearson", "kendall", "spearman"))
```

-   correlation between x and y

-   with incomplete observations, we may want to set use ="complete.obs"

-   based on the argument method, the function cor() computes

    -   Pearson's r (default)

    -   Kendall's $\tau$ (rank based)

    -   Spearman's $\rho$ (rank based)

Example:

```{r}
x<-rnorm(10)
y<-rnorm(10)
cor(x,y)
```

# Correlation test

-   test for correlation between paired samples

```{r,eval=F}
cor.test(x, y, alternative = c("two.sided","less", "greater"), method = c("pearson","kendall", "spearman"), conf.level = 0.95, ...)
```

-   $H_1$ is defined through the argument alternative

-   the argument `method` sets the correlation coefficient

Example:

```{r}
cor.test(x,y)
```
# Excercise 4 Task 5
