---
title: "Exercise 3: Solutions third part"
format: html
toc: false
---
  
```{r, echo=FALSE, eval=TRUE}
library(ggplot2)
```

## Task 3.1

### In this task we use the data set `ChickWeight`. Load the data set from the package `datasets`.

```{r}
data("ChickWeight")
```



## Task 3.2

### Plot the distribution of the variable `weight` using a histogram.

```{r}
ggplot(ChickWeight)+
  geom_histogram(aes(weight))
```




## Task 3.3

### Change the data type of the variable `Time` to a factor.


```{r}
ChickWeight$Time <- as.factor(ChickWeight$Time)
```


## Task 3.4

### Plot a boxplot for each time point.

```{r}
ggplot(ChickWeight, aes(Time, weight))+
    geom_boxplot()
```



## Task 3.5

### Plot the `weight` as boxplots for the different types of `Diet`.

```{r}
ggplot(ChickWeight, aes(Diet, weight))+
    geom_boxplot()
```


## Task 3.6

### Do the same, as before, but now combine the three variables: Use `Time` on the x-axis, `weight` on the y-axis and use Diet as a fill argument.

```{r}
ggplot(ChickWeight, aes(Time, weight, fill=Diet))+
    geom_boxplot()
```


