---
title: "Exercise 5: Solutions first part"
format: html
toc: false
---
  

### Load the data in R.
```{r}
dat <- read.csv("NHANES1.csv")

# make factors
integer_info <- sapply(dat, is.integer)
integer_info[which(names(integer_info) == "age")] <- FALSE  # age should stay an integer
dat[integer_info] <- lapply(dat[integer_info], as.factor)

```

## Task 1.1

### How strong is the relationship between BMI and systolic blood pressure?

```{r}
summary(dat$bmi)
summary(dat$rr_sys)
cor(dat$bmi, dat$rr_sys,use='complete.obs')
cor.test(dat$bmi, dat$rr_sys)
summary(lm(rr_sys~bmi, data = dat))

```


## Task 1.2

### Does the relationship between systolic blood pressure and BMI change when you adjust for age (categorized)? Interpret the coefficients of the resulting model (when you mean-center BMI before fitting the model, you can also interpret the intercept).


```{r}
dat$age_cat <- cut(dat$age, c(0,50,100), right = F)  # cut the age in two groups

summary(lm(rr_sys ~ bmi + age_cat, data = dat))
dat$bmi_c <- dat$bmi - mean(dat$bmi, na.rm=T)
summary(lm(rr_sys ~ bmi_c + age_cat, data=dat))

```


## Task 1.3

### Try to find a better model to predict systolic blood pressure by including more covariates. Select a number of candidate covariates which in your opinion may be related to systolic blood pressure, and then choose a model selection strategy and a criterion/test for comparing models. Describe the model with the best fit according to your search, and interpret the model coefficients.


```{r}

sub_dat <- dat[ ,c('rr_sys', 'bmi', 'male', 'age_cat', 'diab_lft', 'rdyfood_prvmo', 'alc_lft', 'smokstat')]
complete.data <- sub_dat[complete.cases(sub_dat), ]  # complete.cases deletes all rows with at least one NA
complete.data$bmi <- complete.data$bmi - mean(complete.data$bmi)

# full model
mod1 <- lm(rr_sys ~ bmi + male + age_cat + diab_lft + rdyfood_prvmo + alc_lft + smokstat, data=complete.data)
summary(mod1)


# stepwise selection
library(MASS)
mod2 <- stepAIC(mod1)
summary(mod2)

# another potential stepwise selection
mod3 <- lm(formula = rr_sys ~  bmi + male + age_cat + diab_lft + smokstat, data = complete.data)
extractAIC(mod3)

```

