---
title: "Exercise 5: Solutions second part"
format: html
toc: false
---
  

### Load the data in R.
```{r}
data <- read.csv("NHANES1.csv")

# make factors
integer_info <- sapply(data, is.integer)
integer_info[which(names(integer_info) == "age")] <- FALSE  # age should stay an integer
data[integer_info] <- lapply(data[integer_info], as.factor)

# make age category
age.cat <- rep(0, nrow(data))  # assign new variable to the dataframe
age.cat[data$age>=18 & data$age<=34] <- 1  # fill it with values
age.cat[data$age>=35 & data$age<=49] <- 2
age.cat[data$age>=50 & data$age<=64] <- 3
age.cat[data$age>=65 & data$age<=79] <- 4
age.cat[data$age>=80] <- 5
age.cat <- as.factor(age.cat)
data$age.cat <- age.cat
# alternatively use cut() as before. something like age.cat <- cut(data$age, c(0, 18, 35, 50, 65, 80, 100))
```

## Task 2.1

### Analyze the relationship between lifetime diagnosis of cancer and exposure to pollutants, using the categorized age variable (Note: No information on pollutant exposure was collected from participants aged 80+, so these cannot be included in the analysis). Does the adjustment for age change the picture? Interpret the model coefficients including the intercept.

```{r}
# attach the data to use them directly without data$...
# however, use this with caution!
attach(data)

table(workpollut,cancer_ever)
prop.test(table(!workpollut,!cancer_ever))
table(workpollut,cancer_ever,age.cat)
prop.test(table(!workpollut[age.cat==1],!cancer_ever[age.cat==1]))
prop.test(table(!workpollut[age.cat==2],!cancer_ever[age.cat==2]))
prop.test(table(!workpollut[age.cat==3],!cancer_ever[age.cat==3]))
prop.test(table(!workpollut[age.cat==4],!cancer_ever[age.cat==4]))
fisher.test(table(!workpollut,!cancer_ever))
fisher.test(table(!workpollut[age.cat==1],!cancer_ever[age.cat==1]))
fisher.test(table(!workpollut[age.cat==2],!cancer_ever[age.cat==2]))
fisher.test(table(!workpollut[age.cat==3],!cancer_ever[age.cat==3]))
fisher.test(table(!workpollut[age.cat==4],!cancer_ever[age.cat==4]))

mod1 <- glm(cancer_ever ~ workpollut, family = binomial(logit))
summary(mod1)
mod2 <- glm(cancer_ever ~ workpollut + age.cat, family = binomial(logit))
summary(mod2)
```


## Task 2.2

### Try to find a good model of cancer diagnosis, describe and interpret it as you did for systolic blood pressure.


```{r}
sub.data <- data[ ,c('cancer_ever', 'cd', 'pb', 'hg', 'drnkprd_prv12mo', 'cigsprd_prv30d', 'bmi', 'male', 'age.cat', 'diab_lft', 'rdyfood_prvmo')]
complete.data <- sub.data[complete.cases(sub.data), ]

mod3 <- glm(cancer_ever ~ cd + pb + hg + drnkprd_prv12mo + cigsprd_prv30d + bmi + male + age.cat + as.factor(diab_lft) + rdyfood_prvmo, family = binomial(logit), data = complete.data)

library(MASS)
mod4 <- stepAIC(mod3)
summary(mod4)


# explanation
# only intercept
mod5 <- glm(cancer_ever ~ 1, data = data, family=binomial)
summary(mod5)

(tab <- table(data$cancer_ever))
tab["TRUE"]/(tab["TRUE"]+tab["FALSE"])

exp(mod5$coefficients["(Intercept)"])/(1+exp(mod5$coefficients["(Intercept)"]))

# with a predictor
mod6 <- glm(cancer_ever~workpollut, data = data, family = binomial(logit))
summary(mod6)

(tab<-table(data$workpollut, data$cancer_ever))
#For people not exposed to work pollution
p.ne <- tab["FALSE","TRUE"]/(tab["FALSE","TRUE"]+tab["FALSE","FALSE"])
exp(mod6$coefficients["(Intercept)"])/(1+exp(mod6$coefficients["(Intercept)"]))

#For people exposed to work pollution
p.e <- tab["TRUE","TRUE"]/(tab["TRUE","TRUE"]+tab["TRUE","FALSE"])
exp(mod6$coefficients["(Intercept)"]+mod6$coefficients["workpollutTRUE"])/(1+exp(mod6$coefficients["(Intercept)"]+mod6$coefficients["workpollutTRUE"]))

# log-odd baseline (no workpollution)
log(p.ne/(1-p.ne))
mod6$coefficients["(Intercept)"]

# log-odds workpollution
log(p.e/(1-p.e))
mod6$coefficients["(Intercept)"]+mod6$coefficients["workpollutTRUE"]

# log odds-ratio
log((p.e/(1-p.e))/(p.ne/(1-p.ne)))


summary(mod6)
```


