---
title: "Exercise 1: Solutions third part"
format: html
toc: false
---
  

## Task 3.1

### Define a vector `f1` containing 5 arbitrary elements of the type character.
```{r}
f1 <- c("apple", "banana", "cherry", "date", "elderberry")
```


## Task 3.2

### Define a vector `f2` containing 5 arbitrary elements of the type factor.
```{r}
f2 <- factor(c("red", "green", "blue", "yellow", "purple"))
```


## Task 3.3

### Define a vector `f3?? containing 5 other arbitrary elements of the type numeric.
```{r}
f3 <- c(10.5, 23.3, 42.7, 5.8, 12.1)
```


## Task 3.4

### Create a list `L` containing the vectors `f1`, `f2`, `f3`.
```{r}
L <- list(f1 = f1, f2 = f2, f3 = f3)
L
```


## Task 3.5

### Look at the structure of the list.
```{r}
str(L)
```


## Task 3.6

### Create a `data.frame` `df1` using `L`. Look at the structure again.
```{r}
df1 <- data.frame(L)
str(df1)
```
Looks like a list, but with a little bit more info


## Task 3.7

### What are the element on the second row?
```{r}
second_row <- df1[2, ]
second_row
```


## Task 3.8

### What are the element on the second column?
```{r}
second_column <- df1[, 2]
second_column
```


## Task 3.9

### What are the values between the 2nd and the 4th rows?
```{r}
rows_2_to_4 <- df1[2:4, ]
rows_2_to_4
```


## Task 3.10

### Save the data set as a csv.
```{r}
write.csv(df1, file = "data_set.csv", row.names = FALSE)
```


## Task 3.11

### Load the data set into R as a new object `df2` and compare it with the original one `df1`.
```{r}
df2 <- read.csv("data_set.csv")
all.equal(df1, df2)
```

The second variable is not a factor! We can change the type to factor:
```{r}
df2$f2 <- as.factor(df2$f2)
all.equal(df1, df2)
```

