---
title: "Exercise 1: Solutions first part"
format: html
toc: false
---
  

## Task 1.1

### Assign the number 2 to an object `a`.
```{r}
a <- 2
```




## Task 1.2

### Take the natural logarithm of `a`.
```{r}
log_a <- log(a)
log_a
```




## Task 1.3

### Assign  a new object using `sin <- a`.
```{r}
sin <- a
sin
```




## Task 1.4

### What will be the output of `sin(sin)`? Try to predict the output *before* executing the code!
```{r}
sin(sin)
```

No error here as R is smart enough to interpret the oobject with `(...)` as a function!




## Task 1.5

### Assign a new variable `b` with the value `TRUE`.
```{r}
b <- TRUE
b
```




## Task 1.6

### What will be the result of `1 + b`? Why?
```{r}
result_1_plus_b <- 1 + b
result_1_plus_b
```
Explanation: In R, TRUE will b automatically casted to a numeric value of 1. So 1 + 1 = 2.




## Task 1.7

### What will be the result of `sqrt(b)`?
```{r}
result_sqrt_b <- sqrt(b)
result_sqrt_b
```
Explanation: Since TRUE is 1 in numeric terms, sqrt(1) = 1.




## Task 1.8

### What is the value of `b+b`?
```{r}
result_b_plus_b <- b + b
result_b_plus_b
```
Explanation: TRUE + TRUE is 1 + 1 = 2.




## Task 1.9
### Assign a new variable `c` with the value `'1'`.
```{r}
c <- "1"
c
```





## Task 1.10
### What will be the result of `1 + c`? Why?

Explanation: This will result in an error because c is a character string and cannot be added to a numeric value.
```{r}
# Next line will throw an error
# result_1_plus_c <- 1 + c
```
We could use the function `as.numeric` and try to make a numeric from it:
```{r}
result_1_plus_c <- 1 + as.numeric(c)
result_1_plus_c
```



