---
title: "Exercise 2: Solutions"
format: html
toc: false
---
  

## Task 1.1

### Define a vector `x` containing 100 values between 0 and 10.

```{r}
x <- seq(0, 10, lengh.out = 100)
```

## Task 1.2

### Define a second variable `y` using `sin(x)`.

```{r}
y <- sin(x)
```

## Task 1.3

### 

```{r}
plot(x, y, pch = 20)
```

## Task 1.4

### Do the same, but swap the axis now. However, the axis name should stay the same as before.

```{r}
plot(y, x, pch = 20, xlab = "x", ylab = "y")
```

## Task 1.5

### Now, do the same as a line plot. Add a title. Add the additional argument `lty = 2`.

```{r}
plot(x, y, type = "l", lty = 2, main = "Sinus function")

```

## Task 1.6

### Combines a scatter and a line plot: make a scatter plot and use the `lines()` function to add a line.

```{r}
plot(x, y, type = "l", lty = 2, main = "Sinus function")
plot(x, y, pch = 20)
```

## Task 1.7

### Create a new variable `x <- c(1:10, 1:20, 1:30, 1:40, 1:50)`. Plot a histogram using 5 breaks.

```{r}
x = c(1:10, 1:20, 1:30, 1:40, 1:50)
hist(x, breaks = 5)
```

## Task 1.8

### Add a line to the histogram with x-axis values `1:50` and y-axis values `50:1`. Use the additional argument `lwd = 2`.

```{r}
hist(x, breaks = 5)
lines(1:50, 50:1, lwd = 2)
```

## Task 1.9

### Define a data set using the command `df <- data.frame(x = c(rnorm(100), rexp(100)), group = rep(1:2, each = 100))`.

```{r}
df <- data.frame(x = c(rnorm(100), rexp(100)), group = rep(1:2, each = 100))
```

## Task 1.10

### Make a Boxplot of the variable `x`.

```{r}
boxplot(x)
```

## Task 1.11

### Make a Boxplot for both variables.

```{r}
boxplot(x~group, data = df)
```
