---
title: "Exercise 4: Solutions first part"
format: html
toc: false
---
  


## Task 1.1

### Load the data in R.
```{r}
dat <- read.csv("NHANES1.csv")

```


## Task 1.2

### What is the dimension of the data set? How many rows (samples), and how many columns (variables) does the data set contain? What are the variable names of the data set?
```{r}
dim(dat) # dimension of the data set, first number: number of rows, second number: number of columns
# Alternatively:
nrow(dat)     # number of rows
ncol(dat)     # number of columns

names(dat)    # variable names
colnames(dat) # works too
```



## Task 1.3

### All the variables in the data set are either of a class integer, numeric or boolean (i.e., logical). However, some of the variables should be factors rather than numerical variables. Which ones?
```{r}
str(dat)
```



## Task 1.4

### Change the class of these variables to factor.
```{r}
dat$srhgnrl <- as.factor(dat$srhgnrl)
dat$srphbad_prv30d <- as.factor(dat$srphbad_prv30d)
dat$srmhbad_prv30d <- as.factor(dat$srmhbad_prv30d)
dat$adlimp_prv30d <- as.factor(dat$adlimp_prv30d)
dat$educ <- as.factor(dat$educ)
dat$martlst <- as.factor(dat$martlst)
dat$ethnic <- as.factor(dat$ethnic)
dat$increl <- as.factor(dat$increl)
dat$diab_lft <- as.factor(dat$diab_lft)
dat$jobstat_lwk <- as.factor(dat$jobstat_lwk)
dat$alc_lft <- as.factor(dat$alc_lft)
dat$smokstat <- as.factor(dat$smokstat)
dat$milk_month <- as.factor(dat$milk_month)
```



## Task 1.4

### How many women and how many men are there in your data set?
```{r}
sum(dat$male)
sum(!dat$male)

# Alternatively: 
table(dat$male)
```



## Task 1.5

### What is the mean BMI in the overall population? What is the mean BMI for men and women?
```{r}
mean(dat$bmi, na.rm=T)
```



## Task 1.6

### Who has an higher mercury level in blood: men or women? People with chronic bronchitis or people without it? 'Hispanic', 'White', 'Black' or 'Other/Mixed' people?
```{r}
mean(dat$bmi[dat$male], na.rm=T) # mean bmi for men
mean(dat$bmi[!dat$male], na.rm=T) # # mean bmi for women
```



## Task 1.7

### Use the function summary to get the summarized information on all the variables in the data set.
```{r}
summary(dat)
```


