---
title: "Exercise 4: Solutions second part"
format: html
toc: false
---
  


### Load the data in R.
```{r}
dat <- read.csv("NHANES1.csv")

# make factors
integer_info <- sapply(dat, is.integer)
integer_info[which(names(integer_info) == "age")] <- FALSE  # age should stay an integer
dat[integer_info] <- lapply(dat[integer_info], as.factor)

```



## Task 2.1

### Plot the variable ```rr_sys``` as a function of ```bmi```.
```{r}
library(ggplot2)
ggplot(dat, aes(rr_sys, bmi))+
  geom_point()

# baseplot
plot(dat$rr_sys, dat$bmi)
```




## Task 2.2

### Now we want to plot the variable ```rr_sys```  against ```diab_lft```. Which plot should we use here?
```{r}
ggplot(dat, aes(diab_lft, rr_sys))+
  geom_boxplot()

# baseplot
boxplot(rr_sys~diab_lft, names=c("None", "Prediabetes", "Diabetes"), data = dat)
```



## Task 2.3

### Plot the BMI against ```educ``` and give a short interpretation.
```{r}

ggplot(dat, aes(educ, bmi))+
  geom_boxplot()

# baseplot
boxplot(bmi~educ, data = dat)
```



## Task 2.4

### Plot the histogram of the high-density lipoprotein (HDL) cholesterol levels. How does the distribution of HDL look like?
```{r}
ggplot(dat, aes(hdl))+
  geom_histogram(bins=50)

# baseplot
hist(dat$hdl, breaks=50)

```



## Task 2.5

### Can you convert the variable HDL so that its distribution looks more normal? Create such a variable and add it to your data set.
```{r}
dat$logHDL <- log(dat$hdl)

ggplot(dat, aes(logHDL))+
  geom_histogram(bins=50)

hist(dat$logHDL, breaks=50, main="Distribution of log(HDL)", xlab="")

```





