---
title: "Exercise 4: Solutions 5th part"
format: html
toc: false
---
  


### Load the data in R.
```{r}
dat <- read.csv("NHANES1.csv")

# make factors
integer_info <- sapply(dat, is.integer)
integer_info[which(names(integer_info) == "age")] <- FALSE  # age should stay an integer
dat[integer_info] <- lapply(dat[integer_info], as.factor)

```




## Task 4.1

### Use a chi-square test in order to test whether the presence of chronic bronchitis and the current smoking status are independent.
```{r}
table(dat$cbronch_now, dat$smokstat)
chisq.test(dat$cbronch_now, dat$smokstat)
```




## Task 4.2

### Use a Fisher test to verify the independence between sex and the presence of any kind of liver disease.
```{r}
table(dat$livdis_now, dat$male)
fisher.test(dat$livdis_now, dat$male)
```




## Task 4.3

### Perform a sign test both on hdl and on log-hdl to test the hypothesis that the median of the cholesterol level is 1.30. Is the median significantly different from 1.30? Do you obtain the same results using hdl and logHdl?
```{r}
library(BSDA)
SIGN.test(dat$hdl, md = 1.30)
SIGN.test(log(dat$hdl), md = log(1.30))
```




## Task 4.4

### Use a Mann-Whitney test to test the null hypothesis $H_0 : male weight = female weight$.
```{r}
library(coin)
wilcox_test(weight ~ as.factor(male), data = dat)
```




## Task 4.5

### It has been shown that there is a 'social gradient' in health such that the richer you are, the more likely you are to have better health. Plot general self-rated health against relative income so that you can get an impression whether this is confirmed by our data. Which kind of plot is reasonable? Consider using a mosaic plot. E.g. function `mosaicplot()`.
```{r}
mosaicplot(srhgnrl ~ increl,main = 'self-rated health vs. relative income', xlab = 'self-rated health', ylab = 'relative income', data = dat)
```




## Task 4.6

### Test the relation for statistical significance using an appropriate test.

```{r}
chisq.test(dat$srhgnrl, dat$increl)
```




## Task 4.7

### Categorize the variable `bmi` into an underweight (BMI<18.5), normal weight (18.5≤BMI<25), overweight (25≤BMI<30) and obese (BMI≥30) group. Turn the variable into a factor.
```{r}

```
line



## Task 4.8

### What is the proportion of overweight or obese people according to the categorized BMI? What is the proportion of people ever diagnosed with being overweight (variable `ovrwght_ever`)? How many overweight people were actually ever diagnosed with being overweight? 

```{r}
bmi_cat <- NULL
bmi_cat[dat$bmi<18.5] <- "Underweight"
bmi_cat[dat$bmi>=18.5&dat$bmi<25] <- "Normal weight"
bmi_cat[dat$bmi>=25&dat$bmi<30] <- "Overweight"
bmi_cat[dat$bmi>=30] <- "Obese"
bmi_cat <- as.factor(bmi_cat)
dat$bmi_cat <- bmi_cat
prop.table(table(dat$ovrwght_ever))

prop.table(table(dat$ovrwght_ever, dat$bmi_cat), margin = 2)
```




## Task 4.9

### Is there a difference in diabetes prevalence between obese people diagnosed with overweight and those who were never diagnosed? What about self-rated health? How do you explain the results?

```{r}
(tab <- table(dat$diab_lft[dat$bmi_cat=="Obese"], dat$ovrwght_ever[bmi_cat=="Obese"]))
fisher.test(tab)

(tab <- table(dat$srhgnrl[bmi_cat=="Obese"], dat$ovrwght_ever[bmi_cat=="Obese"]))
prop.table(tab, margin=1)
# Chi-square trend test
chisq.test(tab)
```


