---
title: "Exercise 4: Solutions fourth part"
format: html
toc: false
---
  


### Load the data in R.
```{r}
dat <- read.csv("NHANES1.csv")

# make factors
integer_info <- sapply(dat, is.integer)
integer_info[which(names(integer_info) == "age")] <- FALSE  # age should stay an integer
dat[integer_info] <- lapply(dat[integer_info], as.factor)

```



## Task 4.1

### Is the variance the same for males and females? Perform an appropriate test. What is the null hypothesis?
```{r}
var.test(dat$height[dat$male], dat$height[!dat$male])
```




## Task 4.2

### Which test can we use, if, instead, we want to check if males are on average taller than females? Set an adequate alternative hypothesis.
```{r}
t.test(dat$height[dat$male], dat$height[!dat$male], alternative='greater')
```




## Task 4.3

### Analyze the confidence interval obtained in the previous point. Why doesn't it have an upper bound?
```{r}
ci <- t.test(dat$height[dat$male], dat$height[!dat$male], alternative='greater')
ci$conf.int
```




## Task 4.4

### Now look at the distribution of the variable weight: can we graphically state its normality? Perform a transformation in order to recover it.
```{r}
hist(dat$weight,breaks=50)
qqnorm(dat$weight)
qqline(dat$weight)
hist(log(dat$weight),breaks=50)
qqnorm(log(dat$weight))
qqline(log(dat$weight))
```



## Task 4.5

### Test if the mean of the variable ```weight``` is 80 kg, testing $H_0: log(weight) = log(80)$ versus $H_1: log(weight) \neq log(80)$. Can you state an interval estimate with a level of 0.99?
```{r}
t.test(log(dat$weight),mu=log(80))
```





## Task 4.6

### Provide a punctual and an interval estimate for the prevalence of heart diseases and lung pathology.
```{r}
prop.test(table(!dat$heartdis_ever))
prop.test(table(!dat$lungpath_ever))
```




## Task 4.7

### Test if the prevalence is statistically different between men and women.
```{r}
table(dat$male,dat$heartdis_ever)
prop.test(table(dat$male,!dat$heartdis_ever))
chisq.test(table(dat$male, dat$heartdis_ever))
```




## Task 4.8

### Turn the variable smokstat into a binary variable (put the non-smokers and the people who almost never smoked into one group).
```{r}
table(dat$smokstat)
smoke_bin <- NULL
smoke_bin[dat$smokstat == 1 | dat$smokstat == 2] <- 0
smoke_bin[dat$smokstat == 3] <- 1
smoke_bin <- as.factor(smoke_bin)
dat$smoke_bin <- smoke_bin
table(smoke_bin)

```




## Task 4.9

### Do you expect smokers to have a higher cancer prevalence than non-smokers? Test if there is a significant difference in cancer prevalence between smokers and non-smokers (using the variable from before). If so, which group has a higher prevalence? Did you get the result you expected?
```{r}
(tab <- table(dat$smoke_bin, dat$cancer_ever))
prop.table(tab, margin = 1)  # normalize by first margins to 1
prop.test(table(dat$smoke_bin, !dat$cancer_ever))

```




## Task 4.10

### Do the same test in the subgroup of people who are between 20 and 49 years old. What do you see now? How can you explain the results?
```{r}
table(dat$smoke_bin[dat$age < 50], !dat$cancer_ever[dat$age < 50])
prop.test(table(dat$smoke_bin[dat$age < 50], !dat$cancer_ever[dat$age < 50]))

```






